---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(ggplot2)
library(tidyverse)
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

```{r}
av %>% 
  select(
    Name.Alias,
    starts_with("Death")
  ) %>% 
  head()

deaths <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>% 
  select(
    URL, Name.Alias, Time, Died
  )
#head(deaths)
View(deaths)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
?parse_number
deaths$Time <- parse_number(deaths$Time, na = c("", "NA"))
View(deaths)
```

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 
```{r}
#?round
#?count
#?filter
num1 <- deaths %>% group_by(Died) %>% filter(Died == "YES") %>% summarise(total = n()) %>% summarise(total_num = sum(total))
num1

num2 <- deaths %>% group_by(Died) %>% filter(Died != "") %>% summarise(total2 = n()) %>% summarise(total_num2 = sum(total2))
num1 / num2
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

Cole Flickinger's Quote: "I counted 89 total deaths": This is fact checked to be true from the data given.
```{r}
numOfDeaths <- deaths %>% group_by(Died) %>% filter(Died == "YES") %>% summarise(total = n()) %>% summarise(numOfDeaths = sum(total))
numOfDeaths
```

"The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times."
```{r}
jocasta_data <- av %>% 
  filter(str_detect(Name.Alias, "Jocasta")) %>% 
  select(Name.Alias, starts_with("Death"), starts_with("Return"))

jocasta_deaths <- sum(jocasta_data %>% select(starts_with("Death")) == "YES")

jocasta_returns <- sum(jocasta_data %>% select(starts_with("Return")) == "YES")

tibble(
  Character = "Jocasta",
  Deaths = jocasta_deaths,
  Returns = jocasta_returns,
  Perfect_Recovery = (jocasta_deaths == jocasta_returns)
)
```

From the data we can clearly see that Jocasta has died and returned 5 times.


Sofi Gutierrez Quote: "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team"
```{r}
# Count the total Avengers listed in the dataset
total_avengers <- av %>% 
  summarise(total_listed = n())

# Filter Avengers who have died at least once
avengers_with_deaths <- av %>% 
  filter(if_any(starts_with("Death"), ~ .x == "YES")) %>% 
  summarise(total_died = n())

# Display results and validate against the claim
tibble(
  Total_Listed_Avengers = total_avengers$total_listed,
  Avengers_Who_Died_At_Least_Once = avengers_with_deaths$total_died,
  Match_Claim = (total_avengers$total_listed == 173 && avengers_with_deaths$total_died == 69)
)
```
This showing the statement is true
### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
